---
title: "20295 Microeconometrics - Problem Set 2"
author: "Florens Schlosser"
format: html
knitr:
  opts_knit:
    root.dir: "C:/Users/flore/OneDrive/Documents/Bocconi/Year 2/Microeconometrics/PS 2"
---

COMMENT: Before running the code, please adjust the root directory above as well as the file path in the code chunk "data prep" to the desired locations.

# Problem Set 2 - Exercise 2

-----------------------------------------------------
COMMENT: Please change the root directory (at the top ) and the data_path in the Data prep R chunk to point towards the desired directory before running the file.

> We will now use causal forests to assess if there’s any evidence of heterogeneous treatment effects of unilateral divorce laws on divorce rates. The original data set used in Wolfers (2006) did not provide a rich set of variables for this analysis, so we’ll use an expanded version based on simulated observations (the data set is provided on Blackboard as expanded data.csv. These will depict a data set where you would have access to county level observations in each of the states of the original sample, including several characteristics of the population in each county. A table with all variables in the updated data set and their description is provided below.

**Setup**

```{r}
#| label: Libraries

# grf package for random forests
library(grf)

# Necessary in case of time changes
library(timechange)

# Necessary in case of conflicts
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)

# Data stuff
library(tidyverse) 
library(tinytex)
library(fastDummies)
library(stargazer)

# plots
library(ggthemes)
library(viridis)
library(gridExtra)

```

## Exercise 2(a)

> Structure your data set accordingly to assess whether the introduction of the unilateral divorce law had an effect on divorce rates for our sample at the county level. Estimate a causal forest using the causal forest command from package grf.

```{r}
#| label: Data prep

# Use "expanded_data.csv"
data_path <- "C:/Users/flore/OneDrive/Documents/Bocconi/Year 2/Microeconometrics/PS 2/files/expanded_data.csv"
df <- read.csv(data_path)

# Rural dummy: 1 if urbanisation=="Rural", 0 else
df$urbanization_dummy <- as.numeric(df$urbanization == "Rural")

# Assign factor labels to each state for identification
df$state_id <- as.numeric(as.factor(df$st))

# Initialise UNILATERAL variable (= 1 if divorce laws were introduced 1969-1973)
df <- df %>%
  mutate(
    treated = ifelse(lfdivlaw >= 1969 & lfdivlaw <= 1973, 1, 0)
    )
```


```{r}
#| label: Calculate first differences

# calculate change in the simulated divorce rate between 1968 and 1978 per state
df <- df %>% 
  arrange(state_id, county_id, year) %>%
  group_by(state_id) %>% 
  mutate(div_rate_diff = div_rate_sim - lag(div_rate_sim)) %>%        
  ungroup()                                         

# extract baseline-year data (1968)
baseline_df <- df %>%
  filter(year == 1968) 

diff1978_df <- df %>%
  filter(year == 1978) %>%
  select(state_id, county_id, div_rate_diff)

# merge 1968 observations with first-diff variable
final_df <- left_join(baseline_df, diff1978_df, by = c("state_id", "county_id"))

head(final_df)
```

From this dataset, we can now collect the variables of interest for the causal forest, which we will call Y, W, and X:

- **Y**: The outcome variable, in our case first diffs of the divorce rate.

- **W**: The treatment variable, a dummy taking value 1 if the divorce law was implemented between 1968 and 1973 and value 0 else.

- **X**: Covariates (selected population and county characteristics)

```{r}
#| label: Causal forest setup

outcome <- final_df$div_rate_diff.y

treat <- final_df$treated

covariates <- subset(final_df, 
                     select = c("education_rate",
                            "childcare_availability",
                            "unemployment_rate",
                            "median_income",
                            "urbanization_dummy",
                            "marriage_rate",
                            "religious_adherence",
                            "alcohol_consumption",
                            "domestic_violence_rate",
                            "women_labor_force_participation",
                            "crime_rate",
                            "social_services_spending")
                     )
```

We can now estimate the causal forest

```{r}
#| label: Causal forest implementation

tau.forest <- causal_forest(covariates,
                    outcome, 
                    treat
                    )

cate <- average_treatment_effect(tau.forest, 
                                target.sample = "all")

ci_lower <- cate["estimate"] - 1.96 * cate["std.err"]
ci_upper <- cate["estimate"] + 1.96 * cate["std.err"]
cat("ATE:", cate["estimate"], 
    "\nStd. Error:", cate["std.err"],
    "\n95% CI: [", ci_lower, ", ", ci_upper, "]\n")
```

```{r}
varimp <- variable_importance(tau.forest)
ranked.vars <- order(varimp, decreasing = TRUE)

colnames(covariates)[ranked.vars[1:5]]
```

In the following, we conduct a quick sanity check of the causal forest based on the overlap assumption. For this, we regress treatment assignment on the covariates and look at the resulting distribution of propensity scores. It is important that the distribution of propensity scores is concentrated away from 0 and 1 so that we can reasonably assume that no part of the population is systematically excluded from treatment. 

```{r}
# regress treatment on covariates
propensity.forest = regression_forest(covariates, treat) 
W.hat = predict(propensity.forest)$predictions

hist_df <- data.frame(propensity = W.hat)

ggplot(hist_df, aes(x = propensity)) +
  geom_histogram(
    binwidth   = 0.01,           
    fill       = "darkgreen",    
    color      = "white",        
    alpha      = 0.8             
  ) +
  labs(
    title = "Distribution of propensity scores",
    x     = "Propensity score",
    y     = "Count"
  ) +
  theme_minimal() +              
  theme(
    plot.title   = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title   = element_text(size = 14),
    axis.text    = element_text(size = 12)
  )
ggsave(
  filename = file.path("ps2_ex2_graphs","2(a) distribution of propensity scores.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```

As we do not observe a concentration in the distribution of the propensity scores at 0 or 1, we can consider the whole, unadjusted sample when calculating the CATE.

(a)(i)  **What is the estimated average treatment effect in this instance?**

We estimate a Conditional Averate Treatment Effect of `r cate[1]`. Its variance is `r cate[2]``.

(a)(ii)  **Is it consistent with your answer in exercise 1(c)?**

The estimated positive effect is small with a large standard error. In this sense, the results of this exercise are consistent with our findings in exercise 1(c). In both cases, our analysis does not point towards an increasing effect of unilateral divorce laws on divorce rates (in the first decade after adoption), relative to the group of comparison.

## Exercise 2(b)

> Now make an analysis of the causal forest results regarding potential heterogeneous treatment effects. Check the results on (i) the Best Linear Projection; (ii) the Targeting Operator Characteristic; (iii) Plot the distribution of CATEs throughout the distribution of the variables you believe could drive heterogeneity (if you’ll report heterogeneous treatment effects, include graphs for its drivers).
Explain what is being performed in each point and interpret your output.

### 2(b)(i) The Best Linear Projection

```{r}
blp <- best_linear_projection(tau.forest, covariates[ranked.vars[1:5]])
print(blp)
```

**What is being performed**

`best_linear_projection()` estimates a regression of a random forest's doubly‑robust scores on a (selected) set of covariates. The result is a *least‑squares linear approximation* of the conditional average treatment effect (CATE). Importantly, we can perform hypothesis tests on the coefficients of that regression as the random forest scores are orthogonal to the error estimates.

**Output and interpretation**

At least two covariates (domestic violence rate, religious adherence) significantly impact the CATE. Hence, the treatment effect is not homogeneous across counties.

| Covariate | Coefficient | s.e. | \(p\)‑value | Meaning |
|-----------|-------------|------|-------------|---------|
| Domestic‑violence rate | **+0.0488** | 0.0223 | 0.028 | Counties with a high domestic violence rate tend to have a larger increase in the divorce rate after the law change. |
| Religious adherence | **–0.0158** | 0.0038 | <0.001 | We observe larger increases in the divorce rate after the law change in less religious counties. |
| Female labour‑force participation | +0.0116 | 0.0065 | 0.070 | Some indication of a positive correlation between female labour force participation and treatment effect, however p-value > 0.05. |
| Education rate & Social‑services spending | 0.0026  | 0.0046 | > 0.5 | No significant relationship with treatment effect. |
| Crime rate | 0.0002 | 0.0003 | > 0.4 | No significant relationship with treatment effect. |

### 2(b)(ii) The Targeting Operator Characteristic

```{r}
# split sample by state and create training subsample
samples.by.state <- split(seq_along(final_df$state_id), final_df$state_id)
num.states <- length(samples.by.state)
train <- unlist(samples.by.state[sample(1:num.states, num.states / 2)])

# Training forest (In-bag)
train.forest <- causal_forest(
  covariates[train, ],
  outcome[train],
  treat[train],
  clusters = final_df$state_id[train])

tau.hat.eval <- predict(train.forest,covariates[-train, ])$predictions

# Evaluation forest (Out-of-bag)
eval.forest <- causal_forest(
  covariates[-train, ],
  outcome[-train],
  treat[-train],
  clusters = final_df$state_id[-train])

rate.cate <- rank_average_treatment_effect(eval.forest, tau.hat.eval)
png(filename=file.path("ps2_ex2_graphs","2(b)(ii) TOC curve.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By decreasing estimated CATE")
dev.off()
plot(rate.cate, main = "TOC: By decreasing estimated CATE")
```
```{r}
print(rate.cate)
```

```{r}
#| label: Variable Importance after the split

varimp_as <- variable_importance(eval.forest)
ranked.vars_as <- order(varimp_as, decreasing = TRUE)

colnames(covariates)[ranked.vars_as[1:5]]
```

**What is being performed**

* We start by running a random forest to calculate predicted CATEs by county. Then we rank counties by decreasing ATE and consider a fraction \(r\in(0,1]\) of the counties with the highest ATE. Then, we plot the difference between the ATE of each fraction of counties and the ATE of the whole sample as a function of the fraction. Finally, the **Rank‑Weighted Average Treatment Effect (RATE)** calculates a weighted ATE by ranking. Note that due to its dependence on how we split the sample, the RATE has a standard error.

**Output**

```{r}
# calculate CIs for RATE
ci_lower_rate.cate <- rate.cate$estimate - 1.96 * rate.cate$std.err
ci_upper_rate.cate <- rate.cate$estimate + 1.96 * rate.cate$std.err
cat("RATE:", rate.cate$estimate, 
    "\nStd. Error:", rate.cate$std.err,
    "\n95% CI: [", ci_lower_rate.cate, ", ", ci_upper_rate.cate, "]\n")
```

**Interpretation**

As $0\notin CI$, we conclude that treating only the top-CATE share of sample units would significantly increase the ATE. Hence, we can assume that treatment effects are not homogeneous. The TOC-curve shows that, for example, restricting treatment to the 50% highest-ranked fraction of counties would lead to an increase in the ATE of roughly **0.35p.p.**.


### 2(b)(iii) Distribution of CATEs

> Plot the distribution of CATEs throughout the distribution of the variables you believe could drive heterogeneity (if you’ll report heterogeneous treatment effects, include graphs for its drivers).

#### 2(b)(iii)(1) Religious Adherence

```{r}
driver_s1 <- "religious_adherence"

driver <-  covariates[-train, driver_s1]  

rate.rel.ad <- rank_average_treatment_effect(eval.forest, -1 * driver, subset = !is.na(driver))

png(filename=file.path("ps2_ex2_graphs","2(b)(iii) TOC religious adherence.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By increasing religious adherence")
dev.off()

plot(rate.rel.ad, xlab = "Fraction treated", ylab = "Treatment effect", main = "TOC: By increasing religious adherence")
```

```{r}
#| label: Driver 1: religious adherence

scatter_data <- data.frame(
  driver = (covariates[-train, ]$religious_adherence),
  cate   = tau.hat.eval,
  treat  = treat[-train]
)

scatter_data$treat <- factor(scatter_data$treat,
                             levels = c(0,1),
                             labels = c("Control","Treated"))

ggplot(scatter_data, 
       aes(x = driver, 
           y = cate, 
           colour = treat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess",
              se     = TRUE, 
              aes(fill = treat)) +   
  scale_colour_manual(values = c("darkgreen", "steelblue")) +
  scale_fill_manual(  values = c("darkgreen", "steelblue")) +
  labs(title    = "Religious adherence and estimated CATE",
       x        = "Driver (Religious Adherence)",
       y        = "Estimated CATE",
       colour   = "Treatment Group",
       fill     = "Treatment Group") +
  theme_minimal()
ggsave(
  filename = file.path("ps2_ex2_graphs","2(b)(iii) CATE curve religious adherence.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```

#### 2(b)(iii)(2) Domestic Violence Rate

```{r}
driver_s1 <- "domestic_violence_rate"

driver <-  covariates[-train, driver_s1]  

rate.rel.ad <- rank_average_treatment_effect(eval.forest, -1 * driver, subset = !is.na(driver))

png(filename=file.path("ps2_ex2_graphs","2(b)(iii) TOC domestic violence rate.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By increasing dom. violence rate")
dev.off()

plot(rate.rel.ad, xlab = "Fraction treated", ylab = "Treatment effect", main = "TOC: By increasing dom. violence rate")
```

```{r}
#| label: Driver 2: domestic violence rate

scatter_data <- data.frame(
  driver = (covariates[-train, ]$domestic_violence_rate),
  cate   = tau.hat.eval,
  treat  = treat[-train]
)

scatter_data$treat <- factor(scatter_data$treat,
                             levels = c(0,1),
                             labels = c("Control","Treated"))

ggplot(scatter_data, 
       aes(x = driver, 
           y = cate, 
           colour = treat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess",
              se     = TRUE, 
              aes(fill = treat)) +   
  scale_colour_manual(values = c("darkgreen", "steelblue")) +
  scale_fill_manual(  values = c("darkgreen", "steelblue")) +
  labs(title    = "Domestic violence rate and estimated CATE",
       x        = "Driver (Domestic Violence Rate)",
       y        = "Estimated CATE",
       colour   = "Treatment Group",
       fill     = "Treatment Group") +
  theme_minimal()
ggsave(
  filename = file.path("ps2_ex2_graphs","2(b)(iii) CATE curve domestic violence rate.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```

#### 2(b)(iii)(3) Women Labour Force Participation

```{r}
driver_s1 <- "women_labor_force_participation"

driver <-  covariates[-train, driver_s1]  

rate.rel.ad <- rank_average_treatment_effect(eval.forest, -1 * driver, subset = !is.na(driver))

png(filename=file.path("ps2_ex2_graphs","2(b)(iii) TOC female labour participation.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By increasing female labour participation")
dev.off()

plot(rate.rel.ad, xlab = "Fraction treated", ylab = "Treatment effect", main = "TOC: By incr. female labour participation")
```

```{r}
#| label: Driver 3: women labour force participation

scatter_data <- data.frame(
  driver = (covariates[-train, ]$women_labor_force_participation),
  cate   = tau.hat.eval,
  treat  = treat[-train]
)

scatter_data$treat <- factor(scatter_data$treat,
                             levels = c(0,1),
                             labels = c("Control","Treated"))

ggplot(scatter_data, 
       aes(x = driver, 
           y = cate, 
           colour = treat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess",
              se     = TRUE, 
              aes(fill = treat)) +   
  scale_colour_manual(values = c("darkgreen", "steelblue")) +
  scale_fill_manual(  values = c("darkgreen", "steelblue")) +
  labs(title    = "Female labour participation and estimated CATE",
       x        = "Driver (Women Labour Force Participation)",
       y        = "Estimated CATE",
       colour   = "Treatment Group",
       fill     = "Treatment Group") +
  theme_minimal()
ggsave(
  filename = file.path("ps2_ex2_graphs","2(b)(iii) CATE curve female labour participation.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```

#### 2(b)(iii)(4) Education Rate

```{r}
driver_s1 <- "education_rate"

driver <-  covariates[-train, driver_s1]  

rate.rel.ad <- rank_average_treatment_effect(eval.forest, -1 * driver, subset = !is.na(driver))

png(filename=file.path("ps2_ex2_graphs","2(b)(iii) TOC education rate.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By increasing education rate")
dev.off()

plot(rate.rel.ad, xlab = "Fraction treated", ylab = "Treatment effect", main = "TOC: By increasing education rate")
```

```{r}
#| label: Driver 4: education rate

scatter_data <- data.frame(
  driver = (covariates[-train, ]$education_rate),
  cate   = tau.hat.eval,
  treat  = treat[-train]
)

scatter_data$treat <- factor(scatter_data$treat,
                             levels = c(0,1),
                             labels = c("Control","Treated"))

ggplot(scatter_data, 
       aes(x = driver, 
           y = cate, 
           colour = treat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess",
              se     = TRUE, 
              aes(fill = treat)) +   
  scale_colour_manual(values = c("darkgreen", "steelblue")) +
  scale_fill_manual(  values = c("darkgreen", "steelblue")) +
  labs(title    = "Education rate and estimated CATE",
       x        = "Driver (Education rate)",
       y        = "Estimated CATE",
       colour   = "Treatment Group",
       fill     = "Treatment Group") +
  theme_minimal()
ggsave(
  filename = file.path("ps2_ex2_graphs","2(b)(iii) CATE curve education rate.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```
#### 2(b)(iii)(5) Childcare Availability

```{r}
driver_s1 <- "childcare_availability"

driver <-  covariates[-train, driver_s1]  

rate.rel.ad <- rank_average_treatment_effect(eval.forest, -1 * driver, subset = !is.na(driver))

png(filename=file.path("ps2_ex2_graphs","2(b)(iii) TOC childcare availability.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By increasing childcare availability")
dev.off()

plot(rate.rel.ad, xlab = "Fraction treated", ylab = "Treatment effect", main = "TOC: By increasing childcare availability")
```

```{r}
#| label: Driver 5: childcare availability

scatter_data <- data.frame(
  driver = (covariates[-train, ]$childcare_availability),
  cate   = tau.hat.eval,
  treat  = treat[-train]
)

scatter_data$treat <- factor(scatter_data$treat,
                             levels = c(0,1),
                             labels = c("Control","Treated"))

ggplot(scatter_data, 
       aes(x = driver, 
           y = cate, 
           colour = treat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess",
              se     = TRUE, 
              aes(fill = treat)) +   
  scale_colour_manual(values = c("darkgreen", "steelblue")) +
  scale_fill_manual(  values = c("darkgreen", "steelblue")) +
  labs(title    = "Childcare availability and estimated CATE",
       x        = "Driver (Education rate)",
       y        = "Estimated CATE",
       colour   = "Treatment Group",
       fill     = "Treatment Group") +
  theme_minimal()
ggsave(
  filename = file.path("ps2_ex2_graphs","2(b)(iii) CATE curve childcare availability.png"),
  width = 7,
  height = 5,
  dpi = 300
  )
```

***Interpretations of graphical results***
The below table summarises our insights from the graphical analysis conducted in 2(b). Overall, we observe that treatment effect heterogeneity is mainly driven by differences in religious adherence and in domestic violence rates. A weaker effect attributable to the distribution of female labour participation is possible.

| Driver | Graphical Clues | TOC | Interpretation |
|--------|----------------|--------------------------|---------------|
| Religious Adherence | Strong downward slope. | Negative for high \(q\). | Significant negative effect on ATE. |
| Domestic Violence Rate | Strong upward slope. | Positive for low \(q\). | Significant positive effect on ATE |
| Female Labour‑Force Participation | Mild upward slope. | Small positive TOC. | Possible weaker effect on ATE. |
| Education Rate | No identifiable trend. | TOC close to zero. | No observable effect on ATE. |
| Child‑Care Availability | No identifiable trend. | TOC close to zero. | No observable effect on ATE. |

## Exercise 2(c)

> Discuss your results. Did you find any evidence of heterogeneous treatment effects? Justify your answer based on your output in the previous items.

The analyses in 2(b) show that heterogeneity in treatment effects is likely present. For a more detailed interpretation please refer our explanations in point 2(b).

## Exercise 2(d)

> An important aspect in the implementation of causal forests is the use of ”honest trees”, as explained in section 2.4 of Wager and Athey (2017). Explain this procedure and why it is important for our estimation of CATEs. Rerun your analysis without ”honest trees” by selecting `honesty = FALSE`.

### 2(d)(i) Review of the "Honest Tree" Concept

A causal tree is grown in two steps: First, the algorithm determines where to split the data, meaning how the covariate space should be partitioned. Second, once the partition is fixed, it computes the treatment effect within each final leaf, typically as a difference in average outcomes.

A tree is considered honest if no observation is used for both of those tasks within the same tree. In practice, this is implemented by randomly dividing each bootstrap sample into two equal parts. One half, the split sample, is used exclusively to search over candidate partitions using a CART-type splitting rule. The other half, the estimation sample, is then passed through the completed tree, and its observed outcomes Y_i and treatment assignments W_i are used to estimate the treatment effect in each terminal node.
This procedure is repeated independently for every tree in the forest. As a result, a given observation may be used for splitting in some trees and for estimation in others, but it never serves both purposes within the same tree.

Honesty is central to CATE estimation for several reasons.

First, it supports valid inference. The honesty condition is what allows Wager and Athey to establish asymptotic normality for the forest estimator and consistency of variance estimators such as the infinitesimal jackknife and out-of-bag methods. Without honesty, confidence intervals that are supposed to have 95% coverage can perform much worse.

Second, it reduces bias. When the same observations are used both to choose the splits and to estimate the leaf effects, the tree can overfit random fluctuations in the data. That makes selected leaves appear to have more extreme treatment effects than they really do. Keeping the splitting and estimation stages separate helps recover conditional unbiasedness.

Third, it improves out-of-sample reliability. Honest trees are less prone to overstating differences in treatment effects across subgroups, so the forest tends to produce a more trustworthy ranking of CATEs when applied to new data.


### 2(d)(ii) Comparison of Results

```{r}
#| label: Re-estimation of the causal forest without the honesty property

# Fit the causal forest
tau.forest.hf <- causal_forest(covariates,
                    outcome, 
                    treat,
                    honesty = FALSE
                    )

# Estimate the average treatment effect (ATE)
cate_hf <- average_treatment_effect(tau.forest.hf, 
                                target.sample = "all")
cate_hf

ci_lower <- cate["estimate"] - 1.96 * cate["std.err"]
ci_upper <- cate["estimate"] + 1.96 * cate["std.err"]
cat("ATE:", cate["estimate"], 
    "\nStd. Error:", cate["std.err"],
    "\n95% CI: [", ci_lower, ", ", ci_upper, "]\n")

```

```{r}
blp <- best_linear_projection(tau.forest.hf, covariates[ranked.vars[1:5]])
print(blp)
```

```{r}
samples.by.state <- split(seq_along(final_df$state_id), final_df$state_id)
num.states <- length(samples.by.state)
train <- unlist(samples.by.state[sample(1:num.states, num.states / 2)])

# Training forest (In-bag)
train.forest_hf <- causal_forest(
  covariates[train, ],
  outcome[train],
  treat[train],
  honesty = FALSE,
  clusters = final_df$state_id[train])

tau.hat.eval_hf <- predict(train.forest_hf,covariates[-train, ])$predictions

# Evaluation forest (Out-of-bag)
eval.forest_hf <- causal_forest(
  covariates[-train, ],
  outcome[-train],
  treat[-train],
  honesty = FALSE,
  clusters = final_df$state_id[-train])

rate.cate_hf <- rank_average_treatment_effect(eval.forest_hf, tau.hat.eval_hf)

png(filename=file.path("ps2_ex2_graphs","2(d)(ii) TOC non-honest forest.png"), width = 1600, height = 1200, res = 300)
plot(rate.cate, main = "TOC: By decreasing estimated CATE")
dev.off()

plot(rate.cate_hf, main = "TOC: By decreasing estimated CATE")
```

1.  **Result: Is your average treatment effect the same?**

Our re‑estimated non-honest random forest yields:

| Specification | ATE | Std.err. |
|---------------|-----|-----------|
| **Honest (default)** | 0.0436 | 0.0650 |
| **Non‑honest** | 0.0416 | 0.0649 |

The average treatment effect does not considerably change in the re-estimation, neither in size nor in precision. The 95% CIs essentially coincide.

This "robustness" of the ATE estimates is due to the construction of the ATE itself: it is itself already an average of all unit‑level effects, hence the bias in the random forest estimator arising from individual units under lack of honesty will be largely averaged out when aggregating over all units.

2.  **When would you expect this to not be the case?**

1.  **Large and concentrated heterogeneity.**  If there are small regions in the sample space where treatment effects differ highly, a non‑honest estimation may be strongly biased by those regions if the respective observations end up being selected often.
2.  **Propensity imbalances.**  When treatment assignment is highly uneven, a split rule based on \(Y\) can pick leaves where treated and control units differ systematically in covariates, deteriorating potential selection bias issues. 
3. **Small or noisy data sets.**  When each leaf only has a few observations, the biasing effect of re‑using outcomes for split selection is more powerful.   
4. **Tree structure.**  The deeper the tree and the tinier each leaf, the larger the effect of an individual outcome on both the split decision and the estimates. If the selection of individuals is biased, then this bias will be amplified under such a tree structure.
5. **CATE distribution impact.**  Some analyses require knowledge of the shape of the CATE distribution (targeting, subgroup policy evaluation, hypothesis testing,...). Thus, even when the mean effect is unchanged without honesty, the arising bias in the CATE distribution can make such analyses inaccurate.




